---
title: "Hands-on Exercise 1: Geospatial Data Wrangling with R"
editor: visual
---

## Overview

In this hands-on exercise, I learn how to import and wrangling geospatial data using appropriate R packages.

# I. Geospatial Data Wrangling with R

## Getting Started

The code chunk below install and load sf and tidyverse packages into R environment

```{r}
pacman::p_load(sf, tidyverse)
```

## Importing Geospatial Data

### (1) Importing polygon feature data

```{r}
mpsz <- st_read(dsn = "data/geospatial", 
                  layer = "MP14_SUBZONE_WEB_PL")
```

### (2) Importing polyline feature data in shapefile form

The code chunk below uses st_read() function of sf package to import CyclingPath shapefile into R as line feature data frame.

```{r}
cyclingpath = st_read(dsn = "data/geospatial", 
                         layer = "CyclingPath")
```

### (3) Importing GIS data in kml format

The pre-schools-location-kml is in kml format. The code chunk below will be used to import the kml into R. Notice that in the code chunk below, the complete path and the kml file extension were provided.

```{r}
preschool = st_read("data/geospatial/pre-schools-location-kml.kml")
```

The message above reveals that preschool is a point feature data frame. There are a total of 1359 features and 2 fields. Different from the previous two simple feature data frame, preschool is in wgs84 coordinates system.

## Checking the Content of A Simple Feature Data Frame

In this sub-section, you will learn different ways to retrieve information related to the content of a simple feature data frame.

### (1)Working with st_geometry()

The column in the sf data.frame that contains the geometries is a list, of class sfc. We can retrieve the geometry list-column in this case by mpsz\$geom or mpsz\[\[1\]\], but the more general way uses st_geometry() as shown in the code chunk below.

```{r}
st_geometry(mpsz)
```

### (2) Working with glimpse()

Beside the basic feature information, we also would like to learn more about the associated attribute information in the data frame. This is the time you will find glimpse() of dplyr. very handy as shown in the code chunk below.

```{r}
glimpse(mpsz)
```

### (3) Working with head()

Sometimes we would like to reveal complete information of a feature object, this is the job of head() of Base R

```{r}
head(mpsz, n=5)  
```

## Plotting the Geospatial Data

In geospatial data science, by looking at the feature information is not enough. We are also interested to visualise the geospatial features. This is the time you will find plot() of R Graphic comes in very handy as shown in the code chunk below.

```{r}
plot(mpsz)
```

The default plot of an sf object is a multi-plot of all attributes, up to a reasonable maximum as shown above. We can, however, choose to plot only the geometry by using the code chunk below.

```{r}
plot(st_geometry(mpsz))
```

Alternatively, we can also choose the plot the sf object by using a specific attribute as shown in the code chunk below.

```{r}
plot(mpsz["PLN_AREA_N"])

```

## Working with Projection

Map projection is an important property of a geospatial data. In order to perform geoprocessing using two geospatial data, we need to ensure that both geospatial data are projected using similar coordinate system.

In this section, you will learn how to project a simple feature data frame from one coordinate system to another coordinate system. The technical term of this process is called projection transformation.

### (1)Assigning EPSG code to a simple feature data frame

One of the common issue that can happen during importing geospatial data into R is that the coordinate system of the source data was either missing (such as due to missing .proj for ESRI shapefile) or wrongly assigned during the importing process.

This is an example the coordinate system of mpsz simple feature data frame by using st_crs() of sf package as shown in the code chunk below.

```{r}
st_crs(mpsz)
```

Although mpsz data frame is projected in svy21 but when we read until the end of the print, it indicates that the EPSG is 9001. This is a wrong EPSG code because the correct EPSG code for svy21 should be 3414.

In order to assign the correct EPSG code to mpsz data frame, st_set_crs() of sf package is used as shown in the code chunk below.

```{r}
mpsz3414 <- st_set_crs(mpsz, 3414)
```

```{r}
st_crs(mpsz3414)
```

### (2) Transforming the projection of preschool from wgs84 to svy21

```{r}
preschool3414 <- st_transform(preschool, 
                              crs = 3414)
```

## Importing and Converting An Aspatial Data

### (1)Importing the aspatial data

```{r}
listings <- read_csv("data/aspatial/listings.csv")
```

```{r}
list(listings) 
```

### (2)Creating a simple feature data frame from an aspatial data frame

```{r}
listings_sf <- st_as_sf(listings, 
                       coords = c("longitude", "latitude"),
                       crs=4326) %>%
  st_transform(crs = 3414)
```

```{r}
glimpse(listings_sf)
```

## Geoprocessing with sf package

### (1)Buffering

The solution:

Firstly, st_buffer() of sf package is used to compute the 5-meter buffers around cycling paths

```{r}
buffer_cycling <- st_buffer(cyclingpath, 
                               dist=5, nQuadSegs = 30)
```

```{r}
buffer_cycling$AREA <- st_area(buffer_cycling)
```

```{r}
sum(buffer_cycling$AREA)
```

### (2) Point-in-polygon count

```{r}
mpsz3414$`PreSch Count`<- lengths(st_intersects(mpsz3414, preschool3414))
```

```{r}
summary(mpsz3414$`PreSch Count`)
```

```{r}
top_n(mpsz3414, 1, `PreSch Count`)
```

```{r}
mpsz3414$Area <- mpsz3414 %>%
  st_area()
```

```{r}
mpsz3414 <- mpsz3414 %>%
  mutate(`PreSch Density` = `PreSch Count`/Area * 1000000)
```

## Explorotary Data Analysis (EDA)

```{r}
hist(mpsz3414$`PreSch Density`)
```

```{r}
ggplot(data=mpsz3414, 
       aes(x= as.numeric(`PreSch Density`)))+
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  labs(title = "Are pre-school even distributed in Singapore?",
       subtitle= "There are many planning sub-zones with a single pre-school, on the other hand, \nthere are two planning sub-zones with at least 20 pre-schools",
      x = "Pre-school density (per km sq)",
      y = "Frequency")
```

### DIY

```{r}
library(ggplot2)
library(units)
ggplot(mpsz3414, aes(x = `PreSch Count`,y = `PreSch Density`)) + geom_point()+labs(title="The relationship between Pre-school Density and Pre-school Count",
       x="Pre-school count (per km sq)", y = "PreSch Desnity")+
  theme_classic()  
```

# II. Choropleth Mapping with R

## Getting Started

```{r}
pacman::p_load(sf, tmap, tidyverse)
```

## Importing Data into R

### Importing Geospatial Data into R

```{r}
mpsz <- st_read(dsn = "data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
mpsz
```

### Importing Attribute Data into R

```{r}
popdata <- read_csv("data/aspatial/respopagesextod2011to2020.csv")
```

### Data Preparation

1.  Data Wrangling

    ```{r}
    popdata2020 <- popdata %>%
      filter(Time == 2020) %>%
      group_by(PA, SZ, AG) %>%
      summarise(`POP` = sum(`Pop`)) %>%
      ungroup()%>%
      pivot_wider(names_from=AG, 
                  values_from=POP) %>%
      mutate(YOUNG = rowSums(.[3:6])
             +rowSums(.[12])) %>%
    mutate(`ECONOMY ACTIVE` = rowSums(.[7:11])+
    rowSums(.[13:15]))%>%
    mutate(`AGED`=rowSums(.[16:21])) %>%
    mutate(`TOTAL`=rowSums(.[3:21])) %>%  
    mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)
    /`ECONOMY ACTIVE`) %>%
      select(`PA`, `SZ`, `YOUNG`, 
           `ECONOMY ACTIVE`, `AGED`, 
           `TOTAL`, `DEPENDENCY`)

    ```

2.  Joining the attribute data and geospatial data

```{r}
popdata2020 <- popdata2020 %>%
  mutate_at(.vars = vars(PA, SZ), 
          .funs = funs(toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)
```

```{r}
mpsz_pop2020 <- left_join(mpsz, popdata2020,
                          by = c("SUBZONE_N" = "SZ"))
```

```{r}
write_rds(mpsz_pop2020, "data/rds/mpszpop2020.rds")
```

## Choropleth Mapping Geospatial Data Using tmap

### (1) Plotting a choropleth map quickly by using qtm()

```{r}
tmap_mode("plot")
qtm(mpsz_pop2020, 
    fill = "DEPENDENCY")
```

### (2) Creating a choropleth map by using tmap's elements

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "Dependency ratio") +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

#### 1. Drawing a base map

```{r}
tm_shape(mpsz_pop2020) +
  tm_polygons()
```

#### 2. Drawing a choropleth map using tm_polygons()

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY")
```

#### 3. Drawing a choropleth map using tm_fill() and \*tm_border()\*\*

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY")
```
