---
title: "In-class Exercise 1"
author: "Xu Siyi"
editor: visual
---

## 1 Overview

In this in-class exercise 1, I will finish [Chapter 3 Spatial Weights and Applications](https://r4gdsa.netlify.app/chap03.html), which help me to understand how to define spatial relationship.

## 2 The Study Area and Data

Two data sets will be used in this hands-on exercise, they are:

-   Hunan county boundary layer. This is a geospatial data set in ESRI shapefile format.

-   Hunan_2012.csv: This csv file contains selected Hunan's local development indicators in 2012.

### 2.1 Getting Started

The code chunk below will install and load tidyverse and sf packages.

```{r}
pacman::p_load(sf,tidyverse,spdep)
```

## 3 Importing Geospatial Data

### 3.1 importing Polygon features

The code chunk below uses st_read() of sf package to import Hunan shapefile into R. The imported shapefile will be simple features Object of sf.

```{r}
hunan = st_read(dsn = "data/geospatial", 
                  layer = "Hunan")
```

### 3.2 Import csv file

Next, we will import Hunan_2012.csv into R by using read_csv() of readr package. The output is R dataframe class.

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

### 3.3 Performing relational join

The code chunk below will be used to update the attribute table of *hunan*\'s SpatialPolygonsDataFrame with the attribute fields of *hunan2012* dataframe. This is performed by using *left_join()* of **dplyr** package.

```{r}
hunan <- left_join(hunan,hunan2012)
```
